What We Eval in the Shadows

### 2021-04-09 paper/main.tex
tax

\subsection{Arguments of \eval}

The argument passed to a call to \eval determines what the function will do. As
we have seen in the previous section, there are cases where the argument is a
value, but these are rare. The common cases are that the argument is some form
of expression, either a single variable, a simple expression or a statement.

To categorize the expressions, consider a function $min(e)$ which for a given
expression $e$ returns a normal form that abstracts many incidental details
allowing the reader to focus on the structure of the evaluated code. The
minimization function performs several simplifications steps. It performs
constant folding of arithmetic and string expressions for base operators, e.g.
$min(\c{1+1})=\c{V}$, value simplification, $min(\c{c(1,2,3+2)})=\c{V}$,
variable absorption, $min(\c{x+y})=\c{X}$, function absorption,
$min(\c{g(f(x),h(z))})=\c{F(F(X))}$ and a number of other simplification. There
are 15 normal forms. We detail the interesting cases next.

\newcommand{\EE}[1]{{{\emph{\framebox{#1}}}}}

@@

Table~\ref{tab:minimizedexpressions} gives the most frequent forms; \#sites
(\%sites) is the number (ratio) of call sites receiving arguments of that form;
\#packages is the number of packages with that form; \#operations is the average
number of instructions performed by the interpreter when given that minimized
expression; and \%envir is the percentage of call sites for which the \c{envir}
argument is explicitly passed, and is not \c{parent.frame()}.

\begin{table}[h]\small
\begin{tabular}{|c|r|r|r|r|r|c|}\hline
$min(e)$& \#sites & \%sites & \#packages & \#operations & \%envir & example\\\hline
\c X&\packageMinimizedcallsitesa &\packageMinimizedpropsitesa &\packageMinimizedpackagea &\packageMinimizedoperationsaRnd &\packageMinimizedpercentenvirnonparenta & \c{y+1}\\\hline
\c{F(F(X))} & \packageMinimizedcallsitesb  & \packageMinimizedpropsitesb & \packageMinimizedpackageb  & \packageMinimizedoperationsbRnd & \packageMinimizedpercentenvirnonparentb & \c{gbov( mean(x), a-1)}\\\hline
\c{V}&\packageMinimizedcallsitesc &\packageMinimizedpropsitesc &\packageMinimizedpackagec &\packageMinimizedoperationscRnd &\packageMinimizedpercentenvirnonparentc& \c{c(42,21,0)}\\\hline
\c{F(X)}& \packageMinimizedcallsitesd & \packageMinimizedpropsitesd & \packageMinimizedpackaged & \packageMinimizedoperationsdRnd & \packageMinimizedpercentenvirnonparentd & \c{seq\_len(iters)} \\\hline
\c{model.frame} & \packageMinimizedcallsitese & \packageMinimizedpropsitese & \packageMinimizedpackagee & \packageMinimizedoperationseRnd & \packageMinimizedpercentenvirnonparente & \c{model.frame(formula = Z ~ U)} \\\hline
\c{F()}& \packageMinimizedcallsitesf & \packageMinimizedpropsitesf & \packageMinimizedpackagef & \packageMinimizedoperationsfRnd & \packageMinimizedpercentenvirnonparentf & \c{rgamma(3, 2, n = 10L)} \\\hline
\c{\$}& \packageMinimizedcallsitesg & \packageMinimizedpropsitesg & \packageMinimizedpackageg & \packageMinimizedoperationsgRnd & \packageMinimizedpercentenvirnonparentg & \c{DF\$B} \\\hline
\c{FUN} & \packageMinimizedcallsitesh & \packageMinimizedpropsitesh & \packageMinimizedpackageh & \packageMinimizedoperationshRnd & \packageMinimizedpercentenvirnonparenth & \c{function(x, y) x + 3 * y} \\\hline
\c{<-} & \packageMinimizedcallsitesi  & \packageMinimizedpropsitesi & \packageMinimizedpackagei & \packageMinimizedoperationsiRnd & \packageMinimizedpercentenvirnonparenti & \c{x[1, 2:3, 2:3] <- value}\\\hline
\c{\{\}} & \packageMinimizedcallsitesj & \packageMinimizedpropsitesj & \packageMinimizedpackagej & \packageMinimizedoperationsjRnd & \packageMinimizedpercentenvirnonparentj & \c{\{ write.csv(iris, tf) \textbackslash n file.size(tf) \}} \\\hline
\c{F(F(<-))} & \packageMinimizedcallsitesk & \packageMinimizedpropsitesk & \packageMinimizedpackagek & \packageMinimizedoperationskRnd & \packageMinimizedpercentenvirnonparentk & \c{opt\$h.weights <- rep(1, length(x))} \\\hline
\end{tabular}
\caption{Minimized expressions} \label{tab:minimizedexpressions}
\end{table}

>>>>

\subsection{The arguments of \eval}

The expressions passed to \eval vary widely. In order to categorize them, let us
use a minimization function $min(e)$ which for a given expression $e$ returns a
normal form that abstracts incidental details allowing the reader to focus on
the structure of the evaluated code. The minimization function performs constant
folding of arithmetic and string expressions for base operators, e.g.
$min(\c{1+1})=\c{V}$, value simplification, $min(\c{c(1,2,3+2)})=\c{V}$,
variable absorption, $min(\c{x+y})=\c{X}$, function absorption,
$min(\c{g(f(x),h(z))})=\c{F(F(X))}$ and a number of other simplifications.
Table~\ref{tab:minimizedexpressions} gives the 10 most frequent forms; \#sites
and \%sites are, respectively, the number and ratio of sites receiving arguments
of that form; \#packages is the number of packages with that form; \#operations
is the average number of instructions performed by the interpreter; and \%envir
is the percentage of sites for which the \c{envir} argument is explicitly passed
to \eval and is not \c{parent.frame()}.

\begin{table}[h]\small
\begin{tabular}{|c|r|r|r|r|r|c|}\hline
  $min(e)$& \#sites & \%sites & \#packages & \#operations & \%envir & example\\\hline
%% JAN: simplifying osme mumers and adding %
\c X&\packageMinimizedcallsitesa & 28\% &\packageMinimizedpackagea & 768 &77\% & \c{y+1}\\\hline
\c{F(F(X))} & \packageMinimizedcallsitesb  & 20\% & \packageMinimizedpackageb  & 466K & 55\% & \c{gbov( mean(x), a-1)}\\\hline
\c{V}&\packageMinimizedcallsitesc & 17\% &\packageMinimizedpackagec &\packageMinimizedoperationscRnd & 71\%& \c{c(42,21,0)}\\\hline
\c{F(X)}& \packageMinimizedcallsitesd & 14\% & \packageMinimizedpackaged & \packageMinimizedoperationsdRnd & 75\% & \c{seq\_len(iters)} \\\hline
\c{model.frame} & \packageMinimizedcallsitese & 7\% & \packageMinimizedpackagee & 3K & 21\% & \c{model.frame(formula = Z ~ U)} \\\hline
\c{F()}& \packageMinimizedcallsitesf & 6\% & \packageMinimizedpackagef & \packageMinimizedoperationsfRnd & 73\% & \c{rgamma(3, 2, n = 10L)} \\\hline
\c{\$}& \packageMinimizedcallsitesg & 6\% & \packageMinimizedpackageg & 2K & 78\% & \c{DF\$B} \\\hline
\c{FUN} & \packageMinimizedcallsitesh & 6\% & \packageMinimizedpackageh & \packageMinimizedoperationshRnd & 23\% & \c{function(x, y) x + 3 * y} \\\hline
\c{<-} & \packageMinimizedcallsitesi  & 5\% & \packageMinimizedpackagei & 11K & 73\% & \c{x[1, 2:3, 2:3] <- value}\\\hline
\c{BLOCK} & \packageMinimizedcallsitesj & 4\% & \packageMinimizedpackagej & 440K & 73\% & \c{\{ write.csv(iris, tf) \textbackslash n file.size(tf) \}} \\\hline
%% \c X&\packageMinimizedcallsitesa &\packageMinimizedpropsitesa\% &\packageMinimizedpackagea &\packageMinimizedoperationsaRnd &\packageMinimizedpercentenvirnonparenta\% & \c{y+1}\\\hline
%% \c{F(F(X))} & \packageMinimizedcallsitesb  & \packageMinimizedpropsitesb & \packageMinimizedpackageb  & \packageMinimizedoperationsbRnd & \packageMinimizedpercentenvirnonparentb & \c{gbov( mean(x), a-1)}\\\hline
%% \c{V}&\packageMinimizedcallsitesc &\packageMinimizedpropsitesc &\packageMinimizedpackagec &\packageMinimizedoperationscRnd &\packageMinimizedpercentenvirnonparentc& \c{c(42,21,0)}\\\hline
%% \c{F(X)}& \packageMinimizedcallsitesd & \packageMinimizedpropsitesd & \packageMinimizedpackaged & \packageMinimizedoperationsdRnd & \packageMinimizedpercentenvirnonparentd & \c{seq\_len(iters)} \\\hline
%% \c{model.frame} & \packageMinimizedcallsitese & \packageMinimizedpropsitese & \packageMinimizedpackagee & \packageMinimizedoperationseRnd & \packageMinimizedpercentenvirnonparente & \c{model.frame(formula = Z ~ U)} \\\hline
%% \c{F()}& \packageMinimizedcallsitesf & \packageMinimizedpropsitesf & \packageMinimizedpackagef & \packageMinimizedoperationsfRnd & \packageMinimizedpercentenvirnonparentf & \c{rgamma(3, 2, n = 10L)} \\\hline
%% \c{\$}& \packageMinimizedcallsitesg & \packageMinimizedpropsitesg & \packageMinimizedpackageg & \packageMinimizedoperationsgRnd & \packageMinimizedpercentenvirnonparentg & \c{DF\$B} \\\hline
%% \c{FUN} & \packageMinimizedcallsitesh & \packageMinimizedpropsitesh & \packageMinimizedpackageh & \packageMinimizedoperationshRnd & \packageMinimizedpercentenvirnonparenth & \c{function(x, y) x + 3 * y} \\\hline
%% \c{<-} & \packageMinimizedcallsitesi  & \packageMinimizedpropsitesi & \packageMinimizedpackagei & \packageMinimizedoperationsiRnd & \packageMinimizedpercentenvirnonparenti & \c{x[1, 2:3, 2:3] <- value}\\\hline
%% \c{\{\}} & \packageMinimizedcallsitesj & \packageMinimizedpropsitesj & \packageMinimizedpackagej & \packageMinimizedoperationsjRnd & \packageMinimizedpercentenvirnonparentj & \c{\{ write.csv(iris, tf) \textbackslash n file.size(tf) \}} \\\hline
\end{tabular}
\caption{Minimized expressions} \label{tab:minimizedexpressions}
\end{table}

15 normal forms. We detail the interesting cases next.

\newcommand{\EE}[1]{{{\emph{\framebox{#1}}}}\\[1mm]}

@@
